INVARIANT REPRESENTATIONS FOR NOISY SPEECH RECOGNITION

### 2016-10-21 paper.tex
Use nips style

% Template for ICASSP-2017 paper; to be used with:
%          spconf.sty  - ICASSP/ICIP LaTeX style file, and
%          IEEEbib.bst - IEEE bibliography style file.
% --------------------------------------------------------------------------
\documentclass{article}
\usepackage{spconf,amsmath,graphicx}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{algorithm2e}
\usepackage{pbox}
\usepackage{multirow}

@@

% Title.
% ------
\title{INVARIANT REPRESENTATIONS FOR NOISY SPEECH RECOGNITION}
%
% Single address.
% ---------------
\name{Dmitriy Serdyuk $^{\dagger}$\thanks{Dmitriy Serdyuk performed the work
  while at IBM}, Kartik Audhkhasi$^{\star}$, Phil\'emon Brakel$^{\dagger}$, Bhuvana Ramabhadran$^{\star}$, Yoshua Bengio$^{\dagger}$\thanks{Yoshua Bengio is a CIFAR Fellow.}}
\address{$^{\dagger}$ Universit\'e de Montr\'eal\\
         $^{\star}$ IBM}

\begin{document}
%\ninept
%
\maketitle
%
\begin{abstract}
    Robustness of a speech recognition system is a challenging task and modern
    automatic speech recognition (ASR) systems need to be resilient to different recording conditions,
    microphone type, accents, and noisy recording environment. We mitigate
    this problem in a fashion similar to recent research on image generation using
    Generative Adversarial Networks and domain adaptation ideas extending
    adversarial gradient-based training. A recent work of Ganin et al. proposes to
    use adversarial training procedure for image domain adaptation making an intermediate
    representation to improve the target performance and at the same time
    to deteriorate the domain classifier performance, which is a separate neural
    network taking the intermediate representation as the input.

    This work focuses on investigation of neural architectures which produce
    representations invariant to certain conditions as the noise type for speech
    recognition. Similar
    to the Ganin et al. work we train the intermediate representation to be invariant
    to noise conditions training three neural networks: the encoder, the recognizer, and the
    noise classifier to produce the representation produced by the encoder to decrease
    the performance of the classifier making the representation invariant to noise conditions. We
    experimentally evaluate proposed architectures on the acoustic modelling task
    for speech recognition on a popular benchmark Aurora-4.
\end{abstract}
%
\begin{keywords}
speech recognition, deep neural networks, invariant representations, adversarial training
\end{keywords}
%
\section{Introduction}
\label{sec:intro}
    One of the most challenging aspects of automatic speech recognition (ASR),
    is the mismatch between the training and testing conditions. During
    testing, a system may encounter new recording conditions, microphone types,
    accents and types of background noises. Even if all testing conditions also appeared
    in the train data, their relative distribution is likely to be
    different. For this reason, it's important to develop systems that are
    robust towards changes in recording conditions in general.

@@

\section{Generative Adversarial Networks}
\label{sec:gans}
    One of the most insightful recent advantages in the area of generative models
    is the models of the family of generative adversarial networks (GAN).
    A GAN consists of two networks: generator and discriminator. The generator 
    network $G$ has an
    input of randomly generated vector and is asked to produce an image 
    similar to the images in the training set. The discriminator network $D$
    can either receive a generated image from the generator $G$ or an image
    from the training set, the task of the discriminator is to distinguish
    between a generated image and a ``real'' image taken from the dataset. Thus
    the discriminator is just a classifier network with a sigmoid output
    and can be trained with the gradient methods (see 
    Section~\ref{sec:optimization}). The main idea of the generative adversarial
    training is that it is possible to propagate the gradient of the discriminator
    network further to the generator network staking them on top of each other.
    This leads to an efficient training procedure and GANs achieve the state of the
    art that time on handwritten digit generation MNIST dataset.

    In other words, these two networks are competing with each other the 
    generator is trying to deceive the discriminator network while it tries
    to do its best to recognize if there was a deception. This game theoretical
    similarity gave a name of adversarial networks. The objective of adversarial
    training is following
    \begin{align*}
        \min_G \max_D V(D, G) = \mathbb{E}_{\bm{x} \sim p_{\text{data}}(\bm{x})}[\log D(\bm{x})] + \\
          \mathbb{E}_{\bm{z} \sim p_{\bm{z}}(\bm{z})}[\log (1 - D(G(\bm{z})))].
    \end{align*}

    One of the properties of GAN networks that there is no straightforward way 
    to evaluate the quality of image generation process since the GAN doesn't 
    output the log-likelihood of its samples. The paper estimated the likelihood
    using the Parzen windows estimator.

@@

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\linewidth]{model.pdf}
    \caption{The model consists of three neural networks. The encoder $E$ produces
    the intermediate representation $h$ which used in the recognizer $R$ and 
    in the discriminator $D$. The hidden representation $h$ is trained to improve
    the recognition and minimize the discrimination accuracy. The discriminator
    is a classifier trained to maximize its accuracy on the gender or the noise type
    recognition task.}
    \label{fig:model}
\end{figure}

\section{Related Work}

\section{Experiments}
We experimentally evaluate the approach discussed in Section~\ref{sec:gradient-reverse} 
on a noisy speech recognition dataset Aurora-4~\cite{parihar2002aurora}. This dataset
is based on the Wall Street Journal corpus (WSJ0 is available at LDC). Additive noise of 
six categories (airport, babble, car, restaurant, street, train) was introduced to the 
clean data. We used mel-filterbank features with their deltas and delta-deltas spliced
over 5 frames on each side making 11 consecutive frames and 1320 input features in total. 
We did not use any speaker normalization of input features. We pretrained 5 layer
speech recognition network with 2048 rectified linear units at every layer using Adam 
optimization algorithm for 5 epochs. After pretraining, we turned on the invariance term.

@@

\begin{table}
    \centering
    \caption{Results for word error rate (WER\%) on Aurora-4 dataset.}
    \label{tab:results}
    \begin{tabular}{r|cc|cc}
    \multirow{2}{*}{Subset} & \multicolumn{2}{c}{Orig}  & \multicolumn{2}{|c}{Inv}\\ 
               & wv1  & wv2  & wv1  & wv2 \\
    \hline 
    clean      & 4.04 & 6.76 & & \\
    airport    & 7.44 & 15.36& & \\
    babble     & 7.86 & 15.88& & \\
    car        & 4.17 & 8.99 & & \\
    restaurant & 9.25 & 17.32& & \\
    street     & 7.79 & 16.14& & \\
    train      & 8.18 & 17.26& & \\
    \hline
    Average    & \multicolumn{2}{c}{10.46}& & 
    \end{tabular}
    %TODO: add numbers!
\end{table}

@@

We would like to thank Yaroslav Ganin, David Warde-Farley for insightful discussions,
developers of Theano~\cite{2016arXiv160502688short}, Blocks, and Fuel~\cite{MerrienboerBDSW15} 
for great instruments.

% References should be produced using the bibtex program from suitable
% BiBTeX files (here: strings, refs, manuals). The IEEEbib.bst bibliography
% style file from IEEE produces unsorted bibliography list.
% -------------------------------------------------------------------------
\bibliographystyle{IEEE}
\bibliography{refs}

>>>>

% Template for ICASSP-2017 paper; to be used with:
%          spconf.sty  - ICASSP/ICIP LaTeX style file, and
%          IEEEbib.bst - IEEE bibliography style file.
% --------------------------------------------------------------------------
\documentclass{article}

\usepackage[final]{nips_2016}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[draft]{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

\usepackage{amsmath,graphicx}
\usepackage{bm}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{algorithm2e}
\usepackage{pbox}
\usepackage{multirow}

@@

% Title.
% ------
\title{INVARIANT REPRESENTATIONS FOR NOISY SPEECH RECOGNITION}
%
% Single address.
% ---------------
%\name{Dmitriy Serdyuk $^{\dagger}$\thanks{Dmitriy Serdyuk performed the work
%  while at IBM}, Kartik Audhkhasi$^{\star}$, Phil\'emon Brakel$^{\dagger}$, Bhuvana Ramabhadran$^{\star}$, Yoshua Bengio$^{\dagger}$\thanks{Yoshua Bengio is a CIFAR Fellow.}}
%\address{$^{\dagger}$ Universit\'e de Montr\'eal\\
%         $^{\star}$ IBM}
\author{
  Dmitriy Serdyuk\thanks{Dmitriy Serdyuk performed the work 
    during an internship at IBM} \\
  MILA\\
  Universit\'e de Montr\'eal\\
  Montr\'eal, QC H3T 1J4 \\
  \texttt{serdyuk@iro.umontreal.ca} \\
  \And
  Kartik Audhkhasi \\
  IBM \\
  Yorktown Heights, NY 10598\\
  \texttt{kaudhkha@us.ibm.com} \\
  \And
  Phil\'emon Brakel \\
  MILA\\
  Universit\'e de Montr\'eal\\
  Montr\'eal, QC H3T 1J4 \\
  \texttt{pbpop3@gmail.com} \\
  \And
  Bhuvana Ramabhadran\\
  IBM \\
  Yorktown Heights, NY 10598\\
  \texttt{bhuvana@us.ibm.com} \\
  \And
  Yoshua Bengio\\
  MILA, CIFAR Fellow\\
  Universit\'e de Montr\'eal\\
  Montr\'eal, QC H3T 1J4 \\
  \texttt{findme@inter.net} \\
}

\begin{document}
%\ninept
%
\maketitle
%
\begin{abstract}
    Robustness of a speech recognition system is a challenging task and modern
    automatic speech recognition (ASR) systems need to be resilient to different recording conditions,
    microphone type, accents, and noisy recording environment. We mitigate
    this problem in a fashion similar to recent research on image generation using
    Generative Adversarial Networks and domain adaptation ideas extending
    adversarial gradient-based training. A recent work of Ganin et al. proposes to
    use adversarial training procedure for image domain adaptation making an intermediate
    representation to improve the target performance and at the same time
    to deteriorate the domain classifier performance, which is a separate neural
    network taking the intermediate representation as the input.
    This work focuses on investigation of neural architectures which produce
    representations invariant to certain conditions as the noise type for speech
    recognition. Similar
    to the Ganin et al. work we train the intermediate representation to be invariant
    to noise conditions training three neural networks: the encoder, the recognizer, and the
    noise classifier to produce the representation produced by the encoder to decrease
    the performance of the classifier making the representation invariant to noise conditions. We
    experimentally evaluate proposed architectures on the acoustic modelling task
    for speech recognition on a popular benchmark Aurora-4.
\end{abstract}
%
%\begin{keywords}
%speech recognition, deep neural networks, invariant representations, adversarial training
%\end{keywords}
%
\section{Introduction}
\label{sec:intro}
    One of the most challenging aspects of automatic speech recognition (ASR),
    is the mismatch between the training and testing conditions. During
    testing, a system may encounter new recording conditions, microphone types,
    accents and types of background noises. Even if all testing conditions also appeared
    in the train data, their relative distribution is likely to be
    different. For this reason, it's important to develop systems that are
    robust towards changes in recording conditions in general.

@@

\section{Generative Adversarial Networks}
\label{sec:gans}
    One of the most insightful recent advantages in the area of generative models
    is the models of the family of generative adversarial networks (GAN).
    A GAN consists of two networks: generator and discriminator. The generator 
    network $G$ has an
    input of randomly generated vector and is asked to produce an image 
    similar to the images in the training set. The discriminator network $D$
    can either receive a generated image from the generator $G$ or an image
    from the training set, the task of the discriminator is to distinguish
    between a generated image and a ``real'' image taken from the dataset. Thus
    the discriminator is just a classifier network with a sigmoid output
    and can be trained with the gradient methods. The main idea of the generative adversarial
    training is that it is possible to propagate the gradient of the discriminator
    network further to the generator network stacking them on top of each other.
    This leads to an efficient training procedure and GANs achieved the state of the
    art that time on handwritten digit generation MNIST dataset.

    In other words, these two networks are competing with each other the 
    generator is trying to deceive the discriminator network while it tries
    to do its best to recognize if there was a deception. This game theoretical
    similarity gave a name of adversarial networks. If we put it formally, the 
    objective of adversarial training is
    \begin{align*}
        \min_G \max_D V(D, G) = \mathbb{E}_{\bm{x} \sim p_{\text{data}}(\bm{x})}[\log D(\bm{x})] + 
            \mathbb{E}_{\bm{z} \sim p_{\bm{z}}(\bm{z})}[\log (1 - D(G(\bm{z})))].
    \end{align*}

    %One of the properties of GAN networks that there is no straightforward way 
    %to evaluate the quality of image generation process since the GAN doesn't 
    %output the log-likelihood of its samples. The paper estimated the likelihood
    %using the Parzen windows estimator.

@@

\subsection{Invariant represenations for speech recognition}
\label{sec:invariant-speech}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.3\linewidth]{model.pdf}
    \caption{The model consists of three neural networks. The encoder $E$ produces
    the intermediate representation $h$ which used in the recognizer $R$ and 
    in the discriminator $D$. The hidden representation $h$ is trained to improve
    the recognition and minimize the discrimination accuracy. The discriminator
    is a classifier trained to maximize its accuracy on the gender or the noise type
    recognition task.}
    \label{fig:model}
\end{figure}

\section{Related Work}
    Earlier approaches include various works on denoising including

    In a work~\cite{yusuke2016adversarial} an acoustic model trained in an adversarial fashion is
    applied to an in-house dataset based on Wall Street Journal 1 corrupted with additive noise.

\section{Experiments}
We experimentally evaluate the approach discussed in Section~\ref{sec:gradient-reverse} 
on a noisy speech recognition dataset Aurora-4~\cite{parihar2002aurora}. This dataset
is based on the Wall Street Journal corpus (WSJ0 is available at LDC). Noise of 
six categories (airport, babble, car, restaurant, street, train) was introduced to the 
clean data. This noise was added and filtering simulating sampling frequency was applyed 
both to clean and noisy data. 
We used mel-filterbank features with their deltas and delta-deltas spliced
over 5 frames on each side making 11 consecutive frames and 1320 input features in total. 
Global mean and variance normalization was applied to these features.
We did not use any speaker normalization of input features. We pretrained 5 layer
speech recognition network with 2048 rectified linear units at every layer using Adam 
optimization algorithm for 5 epochs. After pretraining, we turned on the invariance term.

@@

\begin{table}
    \centering
    \caption{Results for word error rate (WER\%) on Aurora-4 dataset.}
    \label{tab:results}
    \begin{tabular}{r|cc|cc}
    \multirow{2}{*}{Subset} & \multicolumn{2}{c}{Orig}  & \multicolumn{2}{|c}{Inv}\\ 
               & wv1  & wv2  & wv1  & wv2 \\
    \hline 
    clean      & 4.04 & 6.76 & 5.03& 7.53 \\
    airport    & 7.44 & 15.36& 8.22& 16.03\\
    babble     & 7.86 & 15.88& 9.06& 17.82\\
    car        & 4.17 & 8.99 & 5.21& 9.81 \\
    restaurant & 9.25 & 17.32& 12.2& 19.60\\
    street     & 7.79 & 16.14& 9.34& 17.56\\
    train      & 8.18 & 17.26& 9.60& 18.66\\
    \hline
    Average    & \multicolumn{2}{c}{10.46}& & 
    \end{tabular}
    %TODO: add numbers!
\end{table}

@@

We would like to thank Yaroslav Ganin, David Warde-Farley, Samuel Thomas for insightful discussions,
developers of Theano~\cite{2016arXiv160502688short}, Blocks, and Fuel~\cite{MerrienboerBDSW15} 
for great instruments.

% References should be produced using the bibtex program from suitable
% BiBTeX files (here: strings, refs, manuals). The IEEEbib.bst bibliography
% style file from IEEE produces unsorted bibliography list.
% -------------------------------------------------------------------------
%\bibliographystyle{IEEE}
\bibliographystyle{authordate1}
\bibliography{refs}
